---
title: "Reproduceable Reasearch: Programming Assignment 1"
output: html_document
---
## Loading in the data and converting the column types
```{r setup,include=FALSE}
library(lubridate)
library(dplyr)
library(ggplot2)
library(Amelia)
```
```{r,echo=TRUE, results='hide'}
## LOAD IN DATA
df <- read.csv("activity.csv")

## MISSING VALUES COUNT PER COLUMN
sapply(df, function(x) sum(is.na(x)))

## CONVERT DATE FORMAT
df$date <- ymd(df$date)
```
### What is the mean total number of steps taken per day?
1. What is the mean number of steps taken per day?
```{r}
df_day <- df %>% 
  group_by(date) %>% 
  summarise(stepsNo = sum(steps))
```
2. Make a histogram
```{r}
df_day %>% 
  ggplot(aes(x=stepsNo)) +
  geom_histogram(binwidth = 2000, fill = "yellow", color = "orange") +
  theme_bw() +
  ggtitle("Histogram to show the distribution of the number of steps taken per day") +
  xlab("Number of steps") +
  ylab("Count") +
  scale_x_continuous(breaks = seq(0,25000,5000)) +
  scale_y_continuous(breaks = seq(0,14,2))
```
3. Calculate the mean and median number of steps taken per day
```{r}
## MEAN
mean(df_day$stepsNo, na.rm = TRUE)

## MEDIAN
median(df_day$stepsNo, na.rm = TRUE)
```
### What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval
```{r}
## TIME SERIES PLOT
df_interval <- df %>% 
  group_by(interval) %>% 
  summarise(steps = mean(steps, na.rm = TRUE)) 
df_interval %>% 
  ggplot(aes(x=interval, y= steps)) +
  geom_line(color="darkgreen") +
  ggtitle("Time series plot to show the average amount of steps taken per time interval") +
  xlab("Interval (5mins)") +
  ylab("Mean number of steps") +
  theme_bw()
```
2. Find maximum number of steps
```{r}
df_interval %>% 
  select(interval, steps) %>% 
  filter(steps == max(steps))

df_interval
```
### Imputing missing values
1. Calculate number of missing values
```{r}
sapply(df, function(x) sum(is.na(x)))

##visualise using amelia
missmap(df)
```
2. Fill in missing values
```{r}
dff <- df
for(x in 1:nrow(dff)){
  if(is.na(dff[x,1])==TRUE){
    dff[x,1] <- df_interval[df_interval$interval %in% dff[x,3],2]
  }
}
missmap(dff)
```
3. Make new histogram
```{r}
df_day <- dff %>% 
  group_by(date) %>% 
  summarise(stepsNo = sum(steps))

df_day %>% 
  ggplot(aes(x=stepsNo)) +
  geom_histogram(binwidth = 2000, fill = "lightgreen", color = "green") +
  theme_bw() +
  ggtitle("Histogram to show the distribution of the number of steps taken per day") +
  xlab("Number of steps") +
  ylab("Count") +
  scale_x_continuous(breaks = seq(0,25000,5000)) +
  scale_y_continuous(breaks = seq(0,14,2))
```
## Are there any differences in activity patterns between weekdays and weekends?
1. Create new factor variable
```{r}
## ADD IN WEEKDAY COLUMN
dff$DayName <- weekdays(as.Date(dff$date))

## ADD IN DAY TYPE COLUMN
dff$DayType <- as.factor(ifelse(dff$DayName %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
```
2. Create time series plot to show difference
```{r}
dff %>% 
  group_by(DayType,interval) %>% 
  summarise(mean_steps = mean(steps)) %>% 
  ggplot(aes(x=interval,y=mean_steps)) +
  geom_line(aes(color=DayType)) +
  facet_grid(rows=vars(DayType)) +
  theme_bw() +
  ggtitle("Time series plots to show the difference between weekends and weekdays") +
  ylab("Mean number of steps taken") +
  xlab("Intervals (5min)")